Unrecognized States: Theory and Cases

### 2020-09-28 teaching.tex
add intro and citations to teaching document

@@

>>>>

\section{Unrecognized states in context (i.e. the strategic situation)}
Unrecognized states destabilize the international system and impoverish their residents. Because the borders of these territories are contested, the threat of violent conflict is ever present. Unrecognized states are unable to sign trade agreements or receive most foreign aid, and most face economic sanctions, a combination that leaves their residents isolated and impoverished. Thus, unrecognized statehood is a profoundly undesirable outcome, and yet it is nonetheless often a stable equilibrium. The long-standing unrecognized states of Somaliland and the Turkish Republic of Northern Cyprus, as well as many of the unrecognized states that emerged when the Soviet Union collapsed, still enjoy de facto independence a quarter century later. Unrecognized states are thus quintessential intractable conflicts: difficult but not impossible to resolve (e.g. Crocker, Hampson, and Aall 2004; 2005).\\

Because unrecognized statehood is such a clearly unfavorable outcome for the two primary parties involved – the unrecognized state itself and the home state from which it is attempting to secede – much of the existing work has either treated unrecognized states as an aberration or a temporary phenomenon, or, conversely, as utterly intractable conflicts rooted in the irrationality of one or more actors. We use game theoretic analysis to challenge these understandings. We argue that unrecognized statehood can, in fact, be a stable equilibrium outcome, and that it can be sustained even when all the players involved are perfectly informed and behaving rationally. Game theory provides value in this context by laying bare the mechanisms by which unrecognized statehood is sustained as a stable equilibrium outcome, and thus illuminating the ways in which these intractable conflicts can be peacefully transformed. \\

We analyze not only the actions of unrecognized states and the home states from which they are attempting to secede, but also the patrons that support these unrecognized states and the actors in the international community who work to induce peaceful settlement. By formalizing the incentives of these third-party actors and the actions available to them, we are able to analyze the conditions under which their conduct can lead to peaceful resolution, and when it can lead to war or continued stalemate (Buzard, Graham, and Horne 2017). \\

We define unrecognized states as territories in which a non-state actor controls territory, governs a population, and seeks but does not receive broad recognition as an independent state. The newest unrecognized states were formed when the Soviet Union collapsed in the early 1990s. In the case of more recent cases of militarily successful secession, such as South Sudan, unrecognized statehood has been avoided. This suggests that, if the stalemates sustaining the six current unrecognized states (Abkhazia, Nagorno-Karabakh, Somaliland, South Ossetia, Transnistria, and the Turkish Republic of Northern Cyprus) can be successfully resolved, a world without unrecognized states is possible (Buzard, Graham and Horne 2019).\\

@@

\section{Works Cited}

Buzard, Kristy, Benjamin AT Graham, and Ben Horne. 2019. Unrecognized States: Theory and Cases and Policy Implications, in Overcoming Intractable Conflicts: New Approaches to Constructive Transformations, Miriam Elman, Catherine Gerard, Galia Golan, and Louis Kriesberg, eds. London: Rowman $\&$ Littlefield.\\

Buzard, Kristy, Benjamin AT Graham, and Ben Horne. 2017. "Unrecognized States: A Theory of Self-Determination and Foreign Influence." The Journal of Law, Economics, and Organization 33 no. 3: 578-611.\\

Crocker, Chester A., Fen Osler Hampson, and Pamela Aall. 2004. Taming intractable conflicts: Mediation in the Hardest Cases. Washington, DC, United States Institute of Peace Press. \\

Crocker, Chester A., Fen Osler Hampson, and Pamela Aall, Ed. 2005. Grasping the Nettle: Analyzing Cases of Intractable Conflicts. Washington, D.C., United States Institute of Peace Press.\\
